# Simulations {#sims}

```{r  include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA) #keeps out the hashtags in the knits
options(scipen=999)#eliminates scientific notation
```

In this section I provide the code used to simulate the data in each of the chapters. You may wish to select one of these scales as the practice option for other lessons.

I have added the code, "eval = FALSE" in each of the chunks. This prevents the chunk from executing when the OER is *building*. If you copy the script you will probably want to remove that code.

## Perceptions of the LGBTQ College Campus Climate Scale

The Perceptions of the LGBTQ College Campus Climate Scale [@szymanski_perceptions_2020] is six items with responses rated on a 7-point Likert scale ranging from 1 (*strongly disagree*) to 7 (*strongly agree*). Higher scores indicate more negative perceptions of the LGBTQ campus climate. Szymanski and Bissonette  have suggested that the psychometric evaluation supports using the scale in its entirety or as subscales composed of the following items:

* College response to LGBTQ students:  
  - My university/college is cold and uncaring toward LGBTQ students. (cold)
  - My university/college is unresponsive to the needs of LGBTQ students. (unresponsive) 
  - My university/college provides a supportive environment for LGBTQ students. ([un]supportive; must be reverse-scored)
* LGBTQ Stigma:  
  - Negative attitudes toward LGBTQ persons are openly expressed on my university/college campus. (negative)
  - Heterosexism, homophobia, biphobia, transphobia, and cissexism are visible on my university/college campus. (heterosexism)
  - LGBTQ students are harassed on my university/college campus. (harassed)

A [preprint](https://www.researchgate.net/publication/332062781_Perceptions_of_the_LGBTQ_College_Campus_Climate_Scale_Development_and_Psychometric_Evaluation/link/5ca0bef945851506d7377da7/download) of the article is available at ResearchGate.Below is the script for simulating item-level data from the factor loadings, means, and sample size presented in the published article. 

### Simulating Item Level Data

This first script simulates "raw" item-level data for the entire scale as if individuals had just taken it. This means that any negatively worded items (i.e., supportive) must be reverse-scored.
```{r eval = FALSE}
set.seed(210827)
SzyT1 <- matrix(c(.88, .73, .73, -.07,-.02, .16, -.03, .10, -.04, .86, .76, .71), ncol=2) #primary factor loadings for the two factors
rownames(SzyT1) <- c("cold", "unresponsive", "supportiveNR", "negative", "heterosexism", "harassed") #variable names for the six items
#rownames(Szyf2) <- paste("V", seq(1:6), sep=" ") #prior code I replaced with above
colnames(SzyT1) <- c("F1", "F2")
SzyCorMat <- SzyT1 %*% t(SzyT1) #create the correlation matrix
diag(SzyCorMat) <- 1
#SzyCorMat #prints the correlation matrix
SzyM <- c(2.31, 3.11, 2.40, 3.18, 4.44, 3.02) #item means
SzySD <- c(1.35, 1.46, 1.26, 1.60, 1.75, 1.50) #item standard deviations; turns out we won't need these since we have a covariance matrix
SzyCovMat <- SzySD %*% t(SzySD) * SzyCorMat #creates a covariance matrix from the correlation matrix
#SzyCovMat #displays the covariance matrix

dfSzyT1 <- as.data.frame(round(MASS::mvrnorm(n=646, mu = SzyM, Sigma = SzyCovMat, empirical = TRUE),0)) #creates the item level data from the sample size, mean, and covariance matrix
dfSzyT1[dfSzyT1>7]<-7 #restricts the upperbound of all variables to be 7 or less
dfSzyT1[dfSzyT1<1]<-1 #resticts the lowerbound of all variable to be 1 or greater
#colMeans(dfSzyT1) #displays column means

library(tidyverse)
library(dplyr)
dfSzyT1 <- dfSzyT1 %>% dplyr::mutate(ID = row_number()) #add ID to each row
dfSzyT1 <- dfSzyT1%>%dplyr::select(ID, everything())#moving the ID number to the first column; requires
dfSzyT1<- dfSzyT1 %>%
  dplyr::mutate(supportive = 8 - supportiveNR) #because the original scale had 1 reversed item, I reversed it so that we can re-reverse it for practice. Remember in reversals we subtract from a number 1 greater than our scaling
dfSzyT1 <- dfSzyT1%>%
  dplyr::select(-supportiveNR)
```

The optional script below will let you save the simulated data to your computing environment as either a .csv file (think "Excel lite") or .rds object (preserves any formatting you might do).
```{r eval = FALSE}
#write the simulated data  as a .csv
write.table(dfSzyT1, file="dfSzyT1.csv", sep=",", col.names=TRUE, row.names=FALSE)
#bring back the simulated dat from a .csv file
dfSzyT1 <- read.csv ("dfSzyT1.csv", header = TRUE)
```

```{r eval = FALSE}
#to save the df as an .rds (think "R object") file on your computer; it should save in the same file as the .rmd file you are working with
saveRDS(dfSzyT1, "dfSzyT1.rds")
#bring back the simulated dat from an .rds file
sdfSzyT1 <- readRDS("dfSzyT1.rds")
```

### Simulating Scale Level Data (for Validity Coefficients)

The following script simulates scale-level data (e.g., total and subscale scores for the Perceptions of Campus Climate Scale) and total scale scores for the following measures used to evaluate the convergent and discriminant validity. More complete descriptions of these measures are in the Szymanski and Bissonette article [-@szymanski_perceptions_2020].

* LGBTQ victimization (Victimization)
* Satisfaction with college (CollSat)
* Intention to persist in college (Persistence)
* Generalized anxiety disorder symptoms (Anxiety)
* Symptoms of depression (Depression)

```{r eval = FALSE}
Szy_mu <- c(3.16, 2.71, 3.61, .11, 5.61, 4.41, 1.45, 1.29)
Szy_sd <- c(1.26, 1.33, 1.51, .23, 1.15, .53, .80, .78)
Szy_r_mat <- matrix(c(1,   .88, .90, .35, -.56, -.27, .25, .24,
                  .88,  1,  .58, .25, -.59, -.29, .17, .18,
                  .90, .58,  1,  .37, -.41, -.19, .27, .24,
                  .35, .25, .37,  1,  -.22, -.04, .23, .21,
                 -.56,-.59, -.41, -.22, 1,   .53, -.29, -.32,
                 -.27, -.29, -.19, -.04, .53, 1, -.22, -.26,
                  .25, .17,  .27,  .23,  -.29, -.22, 1, .76,
                  .24,  .18, .24, .21, -.32,  -.26, .76,  1), ncol = 8)
Szy_cov_mat <- Szy_sd %*% t(Szy_sd) * Szy_r_mat

set.seed(210907)
SzyDF <- round(as.data.frame(MASS::mvrnorm(n = 646, mu=Szy_mu, Sigma=Szy_cov_mat, tol=1e-3, empirical=TRUE)),2) #adding "tol=1e-3" fixed the not positive matrix error
SzyDF <- round(dplyr::rename(SzyDF, CClimate = V1, CResponse = V2, Stigma = V3, Victimization = V4, CollSat = V5, Persistence = V6, Anxiety = V7, Depression = V8),2)
#round(cor(SzyDF),2)
```

The optional script below will let you save the simulated data to your computing environment as either a .csv file (think "Excel lite") or .rds object (preserves any formatting you might do).
```{r eval = FALSE}
#write the simulated data  as a .csv
write.table(SzyDF, file="SzyDF.csv", sep=",", col.names=TRUE, row.names=FALSE)
#bring back the simulated dat from a .csv file
SzyDF <- read.csv ("SzyDF.csv", header = TRUE)
```

```{r eval = FALSE}
#to save the df as an .rds (think "R object") file on your computer; it should save in the same file as the .rmd file you are working with
saveRDS(SzyDF, "SzyDF.rds")
#bring back the simulated dat from an .rds file
SzyDF <- readRDS("SzyDF.rds")
```

## Trans Discrimination Scale 21

### Simulating Item Level Data

```{r eval = FALSE }
set.seed(210927)
TDSmat <- matrix(c(.75, .74, .74, .70, .62, .62, .55, .46, .01, -.06, .10, .03, .07, .04, .08, -.14, .13, .01, -.00, .18, .02,
                   .06, -.18, -.00, .06, .00, .12, .06, -.10, .77, .70, .53, -.04, .01, .11, -.04, .06, .10, -.04, .08, .12, .05,
                   .09, .02, -.06, -.03, .03, .02, -.04, .11, .12, .12, -.10, .83, .74, .61, -.08, .04, .19, -.00, -.02, .02, .21,
                   .01, .04, .03, -.12, .06, .09, -.03, .03, -.03, .09, .05, -.06, .00, .10, .70, .70, .52, .04, .07, -.09, .09,
                   -.02, .02, -.13, .05, .10, -.01, .09, .13, .01, .09, .09, .10, .05, -.05, .02, .09, -.03, .89, .70, .65, .54), ncol=5) #primary factor loadings for the four factors taken from Table 2 of the manuscript
rownames(TDSmat) <- c("Micro1", "Micro2", "Micro3", "Micro4", "Micro5", "Micro6", "Micro7", "Micro8", "Work1", "Work2", "Work3", "Health1", "Health2", "Health3", "Law1", "Law2", "Law3", "Educ1", "Educ2",  "Educ3", "Educ4") #variable names for the items
colnames(TDSmat) <- c("Microaggressions", "Employment", "Healthcare", "LawEnf", "Education") #component (subscale) names
TDSCorMat <- TDSmat %*% t(TDSmat) #create the correlation matrix via some matrix algebra
diag(TDSCorMat) <- 1
#SzyCorMat #prints the correlation matrix
TDS_M <- c(3.18, 2.98, 3.22, 3.53, 2.93, 2.54, 4.38, 2.47, 1.47, 1.49, 1.48, 1.86, 1.80, 2.77, 1.49, 1.30, 1.38, 2.35, 1.69, 2.26, 1.63) #item means from Table 2
TDS_SD <- c(1.80, 1.46, 1.43, 1.44, 1.28, 1.22, 1.46, 1.41, .96, .87, .86, 1.25, 1.23, 1.75, .81, .65, .71, 1.39, 1.17, 1.45, .97) #item standard deviations from Table 2
TDSCovMat <- TDS_SD %*% t(TDS_SD) * TDSCorMat #creates a covariance matrix (with more matrix algebra) from the correlation matrix
dfTDS <- as.data.frame(round(MASS::mvrnorm(n=220, mu = TDS_M, Sigma = TDSCovMat, empirical = TRUE),0)) #creates the item level data from the sample size, mean, and covariance matrix; wrapped in commands to round to 0 decimal places and format as a df
dfTDS[dfTDS>6]<-6 #restricts the upperbound of all variables to be 6 or less
dfTDS[dfTDS<1]<-1 #resticts the lowerbound of all variable to be 0 or greater
#colMeans(GRMS) #displays column means

#Below is code if you would like an ID number for each case. Expecially at first, the ID number would just need to be removed, so I will not include it in the original simulation. We will add it later.
#library(tidyverse)
#dfTDS <- dfTDS %>% dplyr::mutate(ID = row_number()) #add ID to each row
#dfTDS <- dfTDS%>%dplyr::select(ID, everything())#moving the ID number to the first column; requires
```

### Simulating Scale Level Data (for Validity Coefficients)

The following script simulates scale-level data (e.g., total and subscale scores for the Perceptions of Campus Climate Scale) and total scale scores for the following measures used to evaluate the convergent and discriminant validity. More complete descriptions of these measures are in the Szymanski and Bissonette article [-@szymanski_perceptions_2020].

* Trans discrimination scale total (TDS21)
* Heterosexist harassment, rejection, and discrimination (HHRDS)
* Proximal minority stressors to include
  - Nondiclosure (Nondisclosure)
  - Negative expectations for the future (NegExp)
  - Internalized transphobia (Intern)
* General psychological disgress (Distress)
* General life stress (Stress)

```{r eval = FALSE}
TDS_mu <- c(2.12, 2.13, 13.98, 23.57, 16.32, 27.97, 8.62)
TDS_sd <- c(.66, .76, 4.87, 6.86, 8.50, 8.49, 3.43)
TDS_r_mat <- matrix(c(1, .83, .19, .32, .13, .29, .26,
                  .83,  1,  .20, .32, .11, .33, .28,
                  .19, .20,  1,  .38, .41, .19, .12,
                  .32, .32, .38,  1,  .46, .36, .36,
                 .13, .11, .41, .46, 1,  .46, .35,
                 .29, .33, .19, .36, .46, 1, .35,
                  .26, .28, .12, .36, .35, .72, 1), ncol = 7)
TDS_cov_mat <- TDS_sd %*% t(TDS_sd) * TDS_r_mat

set.seed(210927)
tdsDF <- round(as.data.frame(MASS::mvrnorm(n = 369, mu=TDS_mu, Sigma=TDS_cov_mat, tol=1e-3, empirical=TRUE)),2) #adding "tol=1e-3" fixed the not positive matrix error
tdsDF <- round(dplyr::rename(tdsDF, TDS21 = V1, HHRDS = V2, Nondisclosure = V3, NegExp = V4, Intern = V5, Distress = V6, Stress = V7),2)
#round(cor(tdsDF),2) #Checking the work
```

## Ableist Microaggression Scale

### Simulating Item Level Data

```{r eval = FALSE }
set.seed(210927)
AMSmat <- matrix(c(.74, .75, .65, .58, .62, .01, .05, -.08, .00, .03, .01, .04, .25, -.06, -.02, .11, .18, .25, .26, .14,
                   -.03, .00, .20, -.07, .15, .71, .52, .47, .02, .04, .00, -.01, .01, -.18, .07, .14, -.17, .05, -.12,.16,
                   .11, -.07, -.03, .20, .03, .00, .07, .15, .91, .85, .64, .56, .42, .04, .04, -.15, .03, .13, .07, .14,
                   -.12, .06, .16, -.01, .02, -.07, .05, .20, -.01, .01, .19, .16, .21, .89, .73, .70, .46, .41, .40, .32), ncol=4) #primary factor loadings for the four factors taken from Table 2 of the manuscript
rownames(AMSmat) <- c("Help1", "Help2", "Help3", "Help4", "Help5", "Min1", "Min2", "Min3", "Pers1", "Pers2", "Pers3", "Pers4", "Pers5", "Oth1", "Oth2", "Oth3", "Oth4", "Oth5","Oth6", "Oth7") #variable names for the items
colnames(AMSmat) <- c("Helplessness", "Minimization", "Personhood", "Otherization") #component (subscale) names
AMSCorMat <- AMSmat %*% t(AMSmat) #create the correlation matrix via some matrix algebra
diag(AMSCorMat) <- 1
#SzyCorMat #prints the correlation matrix
AMS_M <- c(1.95, 1.74, 2.11, 1.61, 2.13, 3.28, 3.02, 2.09, 1.63, 1.43, 1.48, 1.44, 1.71, .89, 1.35, 1.06, 1.39, 1.15, .91, 1.42) #item means from Table 2
AMS_SD <- c(1.54, 1.56, 1.52, 1.61, 1.64, 1.85, 1.54, 1.82, 1.56, 1.51, 1.60, 1.64, 1.55, 1.34, 1.46, 1.50, 1.63, 1.42, 1.29, 1.52) #item standard deviations from Table 2
AMSCovMat <- AMS_SD %*% t(AMS_SD) * AMSCorMat #creates a covariance matrix (with more matrix algebra) from the correlation matrix
dfAMS <- as.data.frame(round(MASS::mvrnorm(n=559, mu = AMS_M, Sigma = AMSCovMat, empirical = TRUE),0)) #creates the item level data from the sample size, mean, and covariance matrix; wrapped in commands to round to 0 decimal places and format as a df
dfAMS[dfAMS>5]<-5 #restricts the upperbound of all variables to be 5 or less
dfAMS[dfAMS<0]<-0 #resticts the lowerbound of all variable to be 0 or greater

#Below is code if you would like an ID number for each case. Expecially at first, the ID number would just need to be removed, so I will not include it in the original simulation. We will add it later.
#library(tidyverse)
#dfAMS <- dfAMS %>% dplyr::mutate(ID = row_number()) #add ID to each row
#dfAMS <- dfAMS %>%dplyr::select(ID, everything())#moving the ID number to the first column; requires
```

```{r eval = FALSE}
#quick check to see if it works
#amsPAF1 <- psych::fa(dfAMS, nfactors = 4, fm = "pa", max.iter = 100, rotate = "varimax")# using raw data and specifying the max number of factors
#amsPAF1
```

### Simulating Scale Level Data (for Validity Coefficients)

```{r eval = FALSE}
AMS_mu <- c(1.96, 2.76, 1.51, 1.17, 1.69, 1.93, 1.26, 1.21)
AMS_sd <- c(1.18, 1.34, 1.28, 1.03, .96, 1.08, .99, .40)
AMS_r_mat <- matrix(c(1, .27, .66, .68, .84, .11, .14, -.08,
                  .27,  1,  .36, .30, .52, .28, .34, -.09,
                  .66, .36,  1,  .76, .89, .22, .25, -.09,
                  .68, .30, .76,  1,  .90, .21, .24, -.15,
                 .84, .52, .89, .90, 1,  .24, .29, .13,
                 .11, .28, .22, .21, .24,  1, .75, -.16,
                  .14, .34, .25, .24, .29, .75,  1, -.16,
                 -.08, -.09, -.09, -.15, -.13, -.16, -.13, 1), ncol = 8)
AMS_cov_mat <- AMS_sd %*% t(AMS_sd) * AMS_r_mat

set.seed(210927)
amsDF <- round(as.data.frame(MASS::mvrnorm(n = 833, mu=AMS_mu, Sigma=AMS_cov_mat, tol=1e-2, empirical=TRUE)),2) #adding "tol=1e-#" fixed the not positive matrix error -- but had to wiggle it around until a number (2) worked
amsDF <- round(dplyr::rename(amsDF, Helplessness = V1, Minimization = V2, Personhood = V3, Otherization = V4, AMS = V5, PSS4 = V6, CESD10 = V7, SDRS5 = V8),2)
#round(cor(tdsDF),2) #Checking the work
```


## Gendered Racial Microaggressions Scale for Black Women (Stress Appraisal)

The GRMS has two scales:  frequency and stress appraisal. This simulation is for the stress appraisal scale.

### Simulating Item Level Data

```{r }
set.seed(210921)
GRMSmat <- matrix(c(.69, .69, .60, .59, .55, .55, .54, .50, .41, .41, .04, -.15, .06, .12, .20, -.01, -.22, -.02, .02, .12, -.09, .06, .19, -.03, -.13,
                  .07, -.07, .00, .07, -.18, .22, .23, -.01, .03, .02, .93, .81, .69, .67, .61, .58, .54, -.04, -.07, -.04, .00, .19, .00, .04, .08,
                  -.08, -.08, 00, .06, .16, -.06, .08, .16, .22, .23, -.04, .01, -.05, -.11, -.16, .25, .16, .59, .55, .54, .54, .51, -.12, .08, .03,
                  -.06, .03, .16, .01, .05, .09, -.08, -.06, .07, -.03, -.08, .18, .03, .06, .06, -.21, .21, .21, .03, -.06, .26, -.14, .70, .69, .68), ncol=4) #primary factor loadings for the four factors taken from the stress appraisal (left hand) factor loadings in Table 1 of the manuscript
rownames(GRMSmat) <- c("Obj1", "Obj2", "Obj3", "Obj4", "Obj5", "Obj6", "Obj7", "Obj8", "Obj9", "Obj10", "Marg1", "Marg2", "Marg3", "Marg4", "Marg5", "Marg6", "Marg7", "Strong1", "Strong2", "Strong3", "Strong4", "Strong5", "Angry1", "Angry2", "Angry3") #variable names for the 25 items
colnames(GRMSmat) <- c("Objectified", "Marginalized", "Strong", "Angry") #component (subscale) names
GRMSCorMat <- GRMSmat %*% t(GRMSmat) #create the correlation matrix via some matrix algebra
diag(GRMSCorMat) <- 1
#SzyCorMat #prints the correlation matrix
GRMS_M <- c(1.78,	1.85,	1.97,	1.93,	2.01,	1.76,	1.91,	2.22,	1.83,	1.88, 2,	3.5,	2.43,	3.44,	2.39,	2.89,	2.7, 1.28,	2.25,	1.45,	1.57,	1.4, 2.02,	2.53,	2.39) #item means; I made these up based on the M and SDs for the factors
GRMS_SD <- c(1.11,	1.23,	0.97,	0.85,	1.19,	1.32,	1.04,	0.98,	1.01,	1.03, 1.01,	0.97,	1.32,	1.24,	1.31,	1.42,	1.2, 0.85,	0.94,	0.78,	1.11,	0.84, 1.14,	1.2,	1.21) #item standard deviations; I made these up based on the M and SDs for the factors
GRMSCovMat <- GRMS_SD %*% t(GRMS_SD) * GRMSCorMat #creates a covariance matrix (with more matrix algebra) from the correlation matrix
dfGRMS <- as.data.frame(round(MASS::mvrnorm(n=259, mu = GRMS_M, Sigma = GRMSCovMat, empirical = TRUE),0)) #creates the item level data from the sample size, mean, and covariance matrix; wrapped in commands to round to 0 decimal places and format as a df
dfGRMS[dfGRMS>5]<-5 #restricts the upperbound of all variables to be 5 or less
dfGRMS[dfGRMS<0]<-0 #resticts the lowerbound of all variable to be 0 or greater
#colMeans(GRMS) #displays column means

#Below is code if you would like an ID number for each case. Expecially at first, the ID number would just need to be removed, so I will not include it in the original simulation. We will add it later.
#library(tidyverse)
#dfGRMS <- dfGRMS %>% dplyr::mutate(ID = row_number()) #add ID to each row
#dfGRMS <- dfGRMS%>%dplyr::select(ID, everything())#moving the ID number to the first column; requires
```

The table of validity coefficients only included the coefficients between the GRMS subscales and scales with the scales used to assess convergent and discriminant validity and not the correlations between the non-GRMS scales. At this time, I do not know how to simulate data without the entire matrix.

## Gendered Racial Microaggressions Scale for Asian American Women (Frequency)

The GRMSAAW has two scales:  frequency and stress appraisal. This simulation is for the frequency scale.

### Simulating Item Level Data

Frequency scaling: 0(*never*), 1 (*rarely*), 2(*sometimes*), 3(*often*), 4(*very frequently*), 5(*always*)


```{r eval = FALSE }
set.seed(210927)
GRMSAAWmat <- matrix(c(.83, .79, .75, .72, .70, .69, .69, .69, .63, -.06, -.01, -.02, .21, -.03, -.04, .02, .05, .17, .05, .01, .00, -.06,
                       .07, -.03, -.06, -.02, .08, -.06, -.01, -.03, .13, .85, .76, .75, .70, .10, -.12, -.06, .01, .06, -.06, -.04, .07, .18,
                       -.11, -.06, .04, .02, -.03, .04, .15, .08, -.03, -.10, .11, .13, -.13, .69, .63, .61, .54, .46, -.05, -.02, .14, .14,
                       .03, .05, -.01, -.06, .04, .08, -.13, .03, .02, .07, .06, -.11, -.02, -.08, .13, .09, -.04, -.03, .90, .79, .62, .51), ncol=4) #primary factor loadings for the four factors taken from Table 2 of the manuscript
rownames(GRMSAAWmat) <- c("AS1", "AS2", "AS3", "AS4", "AS5", "AS6", "AS7", "AS8", "AS9", "AF1", "AF2", "AF3", "AF4", "MI1", "MI2", "MI3", "MI4", "MI5", "AUA1", "AUA2", "AUA3", "AUA4") #variable names for the items
colnames(GRMSAAWmat) <- c("Submissiveness", "Fetishism", "Media", "Appearance") #component (subscale) names
GRMSAAWCorMat <- GRMSAAWmat %*% t(GRMSAAWmat) #create the correlation matrix via some matrix algebra
diag(GRMSAAWCorMat) <- 1
#SzyCorMat #prints the correlation matrix
GRMSAAW_M <- c(2.91,	3.3,	3.45,	2.85,	3.89,	3.11,	3.83,	3.07,	2.88, 3.3, 3.64,	3.21,	3.21, 4.2,	4.8,	4.7,	4.5,	4.89, 4.47,	4.69,	4.47,	4.45) #Means estimated from the information in Table 4. I divided the M by the number of items in each scale then "jittered" the number of values I needed around that mean.
GRMSAAW_SD <- c(1.21,	0.81,	1.34,	1.62,	1.89,	0.93,	1.01,	1.17,	1.22, 1.28,	1.47,	1.45,	1.34, 0.78,	0.93,	0.96,	0.88,	0.91, 1.13,	1.15,	1.11,	1.09) #SDs estimated from the information in Table 4. I divided the SD by the number of items in each scale then "jittered" the number of values I needed around that SD
GRMSAAWCovMat <- GRMSAAW_SD %*% t(GRMSAAW_SD) * GRMSAAWCorMat #creates a covariance matrix (with more matrix algebra) from the correlation matrix
dfGRMSAAW <- as.data.frame(round(MASS::mvrnorm(n=304, mu = GRMSAAW_M, Sigma = GRMSAAWCovMat,  empirical = TRUE),0)) #creates the item level data from the sample size, mean, and covariance matrix; wrapped in commands to round to 0 decimal places and format as a df
dfGRMSAAW[dfGRMSAAW>5]<-5 #restricts the upperbound of all variables to be 5 or less
dfGRMSAAW[dfGRMSAAW<0]<-0 #resticts the lowerbound of all variable to be 0 or greater

#Below is code if you would like an ID number for each case. Expecially at first, the ID number would just need to be removed, so I will not include it in the original simulation. We will add it later.
#library(tidyverse)
#dfGRMSAAW <- dfGRMSAAW %>% dplyr::mutate(ID = row_number()) #add ID to each row
#dfGRMSAAW <- dfGRMSAAW %>%dplyr::select(ID, everything())#moving the ID number to the first column; requires
```

```{r include=FALSE}
#quick check to see if it works
#grmsaawPAF1 <- psych::fa(dfGRMSAAW, nfactors = 4, fm = "pa", max.iter = 100, rotate = "oblimin")# using raw data and specifying the max number of factors
#grmsaawPAF1
```


### Simulating Scale Level Data (for Validity Coefficients)

```{r eval = FALSE}
GRMSAAW_mu <- c(29.26, 13.36, 18.08, 23.09, 83.80, 9.16, 23.40, 16.84, 15.63, 56.49, 49.75, 16.52, 31.44)
GRMSAAW_sd <- c(11.17, 5.54, 4.49, 4.51, 20.81, 2.30, 5.79, 5.85, 3.56, 20.30, 21.51, 6.06, 11.64)
GRMSAAW_r_mat <- matrix(c(1, .00, .00, .00, .00, .28, .24, .46, .16, .4, .29, .15, .13,
                          .00, 1, .00, .0, .00, .02, .05, .11, .07, .34, .27, -.04, .21,
                          .0, .0, 1, .0, .0, .18, .2, .01, -.04, .02, .09, .02, .17,
                          .0, .0, .0, 1, .0, -.02, .08, .31, .36, .15, .08, -.05, -.03,
                          .0, .0, .0, .0, 1, .34, .63, .44, .45, .54, .46, .31, -.06, 
                          .28, .02, .18, -.02, .34, 1, .57, .56, .37, .33, .25, .10, .02,
                          .24, .05, .20, .08, .63, .57, 1, .69, .48, .67, .57, .30, .16,
                          .46, .11, .01, .31, .44, .56, .69, 1, .59, .63, .52, .32, .23,
                          .16, .07, -.04, .36, .45, .37, .48, .59, 1, .46, .31, .11, .07,
                          .4, .34, .02, .15, .54, .33, .67, .63, .46, 1, .83, .3, .14,
                          .29, .27, .09, .08, .46, .25, .57, .52, .31, .83, 1, .3, .2, 
                          .15, -.04, .02, -.05, .31, .10, .30, .32, .11, .3, .3, 1,.18,
                          .13, .21, .17, -.03, -.06, .02, .16, .23, .07, .14, .20, .18, 1), ncol = 13)
GRMSAAW_cov_mat <- GRMSAAW_sd %*% t(GRMSAAW_sd) * GRMSAAW_r_mat

set.seed(210927)
grmsaawDF <- round(as.data.frame(MASS::mvrnorm(n = 260, mu=GRMSAAW_mu, Sigma=GRMSAAW_cov_mat, tol=1e-3, empirical=TRUE)),2) #adding "tol=1e-#" fixed the not positive matrix error -- but had to wiggle it around until a number (2) worked
grmsaawDF <- round(dplyr::rename(grmsaawDF, AS = V1, AF = V2, AUA = V3, MI = V4, General = V5, RMAS_F = V6, RMAS_L = V7, RMAS_I = V8, RMAS_E = V9, SSE_L = V10, SSE_R = V11, PHQ9 = V12, IRAAS = V13),2)
#round(cor(tdsDF),2) #Checking the work
```

```{r include=FALSE}
#quickcheck
#round(cor(grmsaawDF),2)
```

