---
title: "Lewis GRMS Model simulation"
author: "lhbikos"
date: "2024-03-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

An archive of an attempt to simulate all the scales in the Lewis and Neville GRMS article. Because correlation coefficients between all variables are not reported, I did not use this in the lesson.


This lesson's research vignette emerges from Lewis and Neville's Gendered Racial Microaggressions Scale for Black Women [-@lewis_construction_2015]. The article reports on two separate studies that comprised the development, refinement, and psychometric evaluation of two parallel versions (stress appraisal, frequency) of the scale. Below, I simulate data from the final construction of the stress appraisal version as the basis of the lecture. Items were on a 6-point Likert scale ranging from 0 (*not at all stressful*) to 5 (*extremely stressful*).

Lewis and Neville [-@lewis_construction_2015] reported support for a total scale score (25 items) and four subscales.  Below, I list the four subscales, their number of items, and a single example item. At the outset, let me provide a content warning. For those who hold this particular identity (or related identities) the content in the items may be upsetting. In other lessons, I often provide a variable name that gives an indication of the primary content of the item. In the case of the GRMS, I will simply provide an abbreviation of the subscale name and its respective item number. This will allow us to easily inspect the alignment of the item with its intended factor, and hopefully minimize discomfort. If you are not a member of this particular identity, I encourage you to learn about these microaggressions by reading the article in its entirety. Please do not ask members of this group to explain why these microaggressions are harmful or ask if they have encountered them.  The four factors, number of items, and sample item are as follows:

* Assumptions of Beauty and Sexual Objectification
  - 10 items
  - "Objectified me based on physical features."
  - Abbreviated in the simulated data as "Obj#"
* Silenced and Marginalized
  - 7 items
  - "Someone has tried to 'put me in my place.'"
  - Abbreviated in the simulated data as "Marg#"
* Strong Black Woman Stereotype
  - 5 items
  - "I have been told that I am too assertive."
  - Abbreviated in the simulated data as "Str#"
* Angry Black Woman Stereotype
  - 3 items
  - "Someone accused me of being angry when speaking calm."
  - Abbreviated in the simulated data as "Ang#"

Three additional scales were reported in the Lewis and Neville article [-@lewis_construction_2015]. 
Fortunately, I was able to find factor loadings from the original psychometric article or subsequent publications. For multidimensional scales, I assign assign variable names according to the scale to which the item belongs (e.g., Info1). In contrast, when subscales or short unidimensional scales were used, I assigned variable names based on item content (e.g., "blue"). In my own work, I prefer item-level names so that I can quickly see (without having to look up the item names) how the items are behaving. While the focus of this series of chapters is on the Gendered Racial Microaggressions Scale for Black Women scale, this simulated data might be useful to you in one or more of the suggestions for practice (e.g., examining the psychometric characteristics of one or the other scales). The scales, their original citation, and information about how I simulated data for each are listed below. One limitation is that the convergent/discriminant validity correlation matrix in the article only included the correlations between the GRMS scales and the other scales -- it did not provide the correlations between the other measures. While this is a totally reasonable decision (likely owing to space limitations imposed by the journal) it may limit the the remainder of the data for further practice.

* **Racial and Ethnic Microaggressions Scale (REMS; [@nadal_racial_2011])** is a 45-item scale with Likert scaling ranging from 1 (*I did not experience this event*) to 5  (*I experienced this event 7 or more times*). Higher scores represent greater frequency of perceived microaggressions within the past six months. I simulated data at the subscale level. There are six subscales:  Assumptions of Inferiority, Second-Class Citizan and Assumption of Criminality, Microinvalidations, Exoticization and Assumptions of Similarity, Environmental Microaggressions, and Workplace and School Microaggressions. Data were simulated using factor loadings in the source article. 
* **Schedule of Sexist Events (SSE; [@klonoff_schedule_1995])**  is a 20-item scale that ranging from 1 (*the event has never happened to me* to 6 (*the event happened almost all [i.e., more than 70%] of the time*). Higher scores represent greater lifetime frequency of everyday sexist events. I simulated data the subscale level. There are three subscales: Sexist Degradation and Its Consequences, Unfair/Sexist Events at Work/School, and Unfair Treatment in Distant and Close Relationships. Data were simulated using factor loadings from Matteson and Moradi's [-@matteson_examining_2005] replication study. Consistent with Lewis and Neville [-@lewis_construction_2015], I only simulated only the lifetime frequency scores (in the original scale there is a parallel assessment of recent experiences).
* **Mental Health inventory 5** (MHI5; [@veit_structure_1983]) is a 5-item scale with Likert scaling ranging from 1 (*all of the time) to 6 (*none of the time*). In the context of this study, higher scores indicate higher levels of psychological distress. I simulated data using factor loadings from the source article.

Below, I walk through the data simulation. This is not an essential portion of the lesson, but I will lecture it in case you are interested. None of the items are negatively worded (relative to the other items), so there is no need to reverse-score any items.

Simulating the data involved using factor loadings, means, standard deviations, and correlations between the scales. Because the simulation will produce "out-of-bounds" values, the code below rescales the scores into the range of the Likert-type scaling and rounds them to whole values.

```{r tidy=TRUE, tidy.opts=list(width.cutoff=70), warning=FALSE, message=FALSE}
#Entering the intercorrelations, means, and standard deviations from the journal article

LewisGRMS_generating_model <- '
        #measurement model
        Objectification =~ .69*Obj1 + .69*Obj2 + .60*Obj3 + .59*Obj4 + .55*Obj5 + .55*Obj6 + .54*Obj7 + .50*Obj8 + .41*Obj9 + .41*Obj10
        Marginalized =~ .93*Marg1 + .81*Marg2 +.69*Marg3 + .67*Marg4 + .61*Marg5 + .58*Marg6 +.54*Marg7
        Strong =~ .59*Str1 + .55*Str2 + .54*Str3 + .54*Str4 + .51*Str5
        Angry =~ .70*Ang1 + .69*Ang2 + .68*Ang3
        
        REMS_Inf =~ .82*Inf32 + .75*Inf38 + .74*Inf21 + .72*Inf17 + .69*Inf9 + .61*Inf36 + .51*Inf5 + .49*Inf22
        REMS_SClass =~ .81*SClass6 + .81*SClass31 + .74*SClass8 + .74*SClass40 + .72*SClass2 + .65*SClass34 + .55*SClass11
        REMS_mInv =~.84*mInv27 + .84*mInv30 + .80*mInv39 + .72*mInv7 + .62*mInv26 + .61*mInv33 + .53*mInv4 + .47*mInv14 + .47*mInv10
        REMS_Exot =~ .74*Exot3 + .74*Exot29 + .71*Exot45 + .69*Exot35 + .60*Exot42 + .59*Exot23 + .51*Exot13 + .51*Exot20 + .49*Exot43
        REMS_mEnv =~ .84*mEnv37 + .85*mEnv24 + .78*mEnv19 + .70*mEnv28 + .69*mEnv18 + .55*mEnv41 + .55*mEnv12
        REMS_mWorks =~.76*mWork25 + .67*mWork15 + .65*mWork1 + .64*mWork16 + .62*mWork44
      
        SSE_Degradation =~ .80*argument + .78*name + .74*telloff + .73*angry + .73*jokes + .72*advances + .68*disrespect + .57*threatened
        SSE_Work =~ .80*employer + .75*denied + .61*teachers + .55*coworkers + .44*different
        SSE_Relationships =~ .66*man + .66*helpers + .66*neighbors + .63*family + .63*service + .49*strangers + .37*drastic
        
        MHI5 =~ .80*nervous + .82*blue + .75*dumps + .83*happy + .77*calm

        #Means
         Objectification ~ 1.85*1
         Marginalized ~ 2.67*1
         Strong ~ 1.61*1
         Angry ~ 2.29*1
         
         REMS_Inf ~ 2.05*1
         REMS_SClass ~ 1.80*1
         REMS_mInv ~ 2.00*1
         REMS_Exot ~ 1.58*1
         REMS_mEnv ~ 3.34*1
         REMS_mWorks ~1.86*1
         
         SSE_Degradation ~ 2.54*1
         SSE_Work ~ 2.68*1
         SSE_Relationships ~2.27*1
         
         MHI5 ~ 2.49*1

        #Correlations
         Objectification =~ .63*Marginalized
         Objectification =~ .66*Strong
         Objectification =~ .51*Angry
         Objectification =~ .22*REMS_Inf
         Objectification =~ .13*REMS_SClass
         Objectification =~ .29*REMS_mInv
         Objectification =~ .24*REMS_Exot
         Objectification =~ .10*REMS_mEnv
         Objectification =~ .18*REMS_mWorks
         Objectification =~ .63*SSE_Degradation
         Objectification =~ .33*SSE_Work
         Objectification =~ .46*SSE_Relationships
         Objectification =~ .23*MHI5
         
         Marginalized =~ .59*Strong
         Marginalized =~ .62*Angry
         Marginalized =~ .39*REMS_Inf
         Marginalized =~ .26*REMS_SClass
         Marginalized =~ .39*REMS_mInv
         Marginalized =~ .27*REMS_Exot
         Marginalized =~ .07*REMS_mEnv
         Marginalized =~ .49*REMS_mWorks
         Marginalized =~ .40*SSE_Degradation
         Marginalized =~ .47*SSE_Work
         Marginalized =~ .50*SSE_Relationships
         Marginalized =~ .26*MHI5
         
         Strong =~ .61*Angry
         Strong =~ .24*REMS_Inf
         Strong =~ .27*REMS_SClass
         Strong =~ .35*REMS_mInv
         Strong =~ .34*REMS_Exot
         Strong =~ -.08*REMS_mEnv
         Strong =~ .28*REMS_mWorks
         Strong =~ .47*SSE_Degradation
         Strong =~ .50*SSE_Work
         Strong =~ .51*SSE_Relationships
         Strong =~ .23*MHI5
         
         Angry =~ .28*REMS_Inf
         Angry =~ .31*REMS_SClass
         Angry =~ .19*REMS_mInv
         Angry =~ .28*REMS_Exot
         Angry =~ -.06*REMS_mEnv
         Angry =~ .16*REMS_mWorks
         Angry =~ .50*SSE_Degradation
         Angry =~ .35*SSE_Work
         Angry =~ .45*SSE_Relationships
         Angry =~ .28*MHI5
        
        '

set.seed(240304)
dfGRMS <- lavaan::simulateData(model = LewisGRMS_generating_model,
                              model.type = "sem",
                              meanstructure = T,
                              sample.nobs=259,
                              standardized=FALSE)

#used to retrieve column indices used in the rescaling script below
col_index <- as.data.frame(colnames(dfGRMS))

#The code below loops through each column of the dataframe and assigns the scaling accordingly
#Rows 1 thru 26 are the GRMS items
#Rows 27 thru 70 are the REMS items
#Rows 71 thru 90 are the SSE items
#Rows 91 thru 95 are the MHI5 items

for(i in 1:ncol(dfGRMS)){  
  if(i >= 1 & i <= 26){   
    dfGRMS[,i] <- scales::rescale(dfGRMS[,i], c(1, 5))
  }
    if(i >= 27 & i <= 70){   
    dfGRMS[,i] <- scales::rescale(dfGRMS[,i], c(1, 5))
    }
        if(i >= 71 & i <= 90){  
    dfGRMS[,i] <- scales::rescale(dfGRMS[,i], c(1, 6))
        }
  if(i >= 91 & i <= 95){  
    dfGRMS[,i] <- scales::rescale(dfGRMS[,i], c(1, 7))
  }
}

#rounding to integers so that the data resembles that which was collected
library(tidyverse)
dfGRMS <- dfGRMS %>% round(0) 

#quick check of my work
#psych::describe(dfGRMS) 

```


A simulation with just the items of the GRMS (and no correlations with other scales).

```{r tidy=TRUE, tidy.opts=list(width.cutoff=70), warning=FALSE, message=FALSE}
#Entering the intercorrelations, means, and standard deviations from the journal article

LewisGRMS_generating_model <- '
        #measurement model
        Objectification =~ .69*Obj1 + .69*Obj2 + .60*Obj3 + .59*Obj4 + .55*Obj5 + .55*Obj6 + .54*Obj7 + .50*Obj8 + .41*Obj9 + .41*Obj10
        Marginalized =~ .93*Marg1 + .81*Marg2 +.69*Marg3 + .67*Marg4 + .61*Marg5 + .58*Marg6 +.54*Marg7
        Strong =~ .59*Str1 + .55*Str2 + .54*Str3 + .54*Str4 + .51*Str5
        Angry =~ .70*Ang1 + .69*Ang2 + .68*Ang3
        
        #Means
         Objectification ~ 1.85*1
         Marginalized ~ 2.67*1
         Strong ~ 1.61*1
         Angry ~ 2.29*1
         
        #Correlations
         Objectification ~~ .63*Marginalized
         Objectification ~~ .66*Strong
         Objectification ~~ .51*Angry
         
         Marginalized ~~ .59*Strong
         Marginalized ~~ .62*Angry

         Strong ~~ .61*Angry
 
        '

set.seed(240311)
dfGRMS <- lavaan::simulateData(model = LewisGRMS_generating_model,
                              model.type = "sem",
                              meanstructure = T,
                              sample.nobs=259,
                              standardized=FALSE)

#used to retrieve column indices used in the rescaling script below
col_index <- as.data.frame(colnames(dfGRMS))

#The code below loops through each column of the dataframe and assigns the scaling accordingly
#Rows 1 thru 26 are the GRMS items

for(i in 1:ncol(dfGRMS)){  
  if(i >= 1 & i <= 26){   
    dfGRMS[,i] <- scales::rescale(dfGRMS[,i], c(1, 5))
  }
}

#rounding to integers so that the data resembles that which was collected
library(tidyverse)
dfGRMS <- dfGRMS %>% round(0) 

#quick check of my work
#psych::describe(dfGRMS) 

```
